---
title: "Portfolio 2"
author: "Yogesh Kumar"
date: "20 May 2018"
output: 
  html_document:
    keep_md: true
---
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = 'png')
```
## install the required packages and load the required libraries
```{r loading Libraries}
library(readr)
library(ggmap)
library(mapproj)
```
## Loading data using read_csv command
```{r loading csv files}
apr <- read_csv("uber-raw-data-apr14.csv")
may <- read_csv("uber-raw-data-may14.csv")
jun <- read_csv("uber-raw-data-jun14.csv")
jul <- read_csv("uber-raw-data-jul14.csv")
aug <- read_csv("uber-raw-data-aug14.csv")
sep <- read_csv("uber-raw-data-sep14.csv")
```
## Merge the data using rbind 
```{r Csv merging and clean up}
uberData <- rbind(apr, may, jun,jul,aug,sep)
```

## Data cleaning
```{r}
uberData <- uberData[!is.na(uberData$Lat),]
uberData <- uberData[!is.na(uberData$Lon),]
uberData <- uberData[uberData$Lat != 0,]
uberData <- uberData[uberData$Lon != 0,]
uberData <- uberData[!is.na(uberData$'Date/Time'),]
```


## fetching subsets of whole data(jun and july) and data data cleaning to proceed with data visualisation
```{r Csv merging for subset of months and clean up}
uberMinorData <- rbind( jun,jul)
uberMinorData <- uberMinorData[uberMinorData$Lat != 0,]
uberMinorData <- uberMinorData[uberMinorData$Lon != 0,]
uberMinorData <- uberMinorData[!is.na(uberMinorData$Lat),]
uberMinorData <- uberMinorData[!is.na(uberMinorData$Lat),]
uberMinorData <- uberMinorData[!is.na(uberMinorData$'Date/Time'),]
```
##Task 1.1
## calculating statistics from the dataset and visualising the data to show daily trends in pick ups.
## Task 1.1 
Day is extracted from Date/Time colum . The code involves explicit consideration of column value as Date/Time. The original csv is update with a new column known as week day. Generate Frequency of Days and Sort Data accordng to Days. Afterwars, plot the data with Day as x axis and Frequency on Y


```{r extraction of day}
uberData$day <- weekdays(as.Date(uberData$'Date/Time',"%m/%d/%Y"))
```

```{r Plot on frequency of pick up to the week day}
Day_count<-as.data.frame(table(uberData$day))
names(Day_count)<-c("Days","Freq")
ggplot(data=Day_count,aes(x=Days,y=Freq,fill=rownames(Day_count)))+geom_bar(stat="identity",colour = "black")+guides(fill=FALSE)+labs(title = "Frequency by Days ",x="Day",y="Frequency")+ geom_text(aes(label=Freq),vjust=1.6,color="white",size=4)+theme_minimal()
```
##Task 1.1 Statistics 
 step 1: Find data statistics and get minimum and maximum
```{r Statistics of frequency vs week day}
uberDataDay<-factor(uberData$day)
table(uberDataDay)
min(table(uberDataDay))
max(table(uberDataDay))
```
##Task 1.1 Summary
The visualization and statistics above demonstrate the following result. Sunday sees the Lowest with a count of 490180. While, Thursday sees the Highest number of Uber Pickups with 755145 and 

##Task 1.2
## Task 1.2 Plot
### Task 1.2 Steps
Month is extracted from Date/Time colum . The code invlves explicit consideration of column value as Date/Time. The original csv is update with a new column knows as month.Generate Frequency of Month and Sort Data accordng to Month. Afterwards, Plot bargraph with Month on x-axis and Frequency on Y-Axis for Visualization

```{r month extraction}
uberData$month <- months(as.Date(uberData$'Date/Time',"%m/%d/%Y"))
```

```{r Plot on frequency vs Month data}
Month_count<-as.data.frame(table(uberData$month))
names(Month_count)<-c("Month","Freq")
ggplot(data=Month_count,aes(x=Month,y=Freq,fill=rownames(Month_count)))+geom_bar(stat="identity",colour = "black")+guides(fill=FALSE)+labs(title = "Frequency by Month ",x="Month of year ",y="Frequency")+ geom_text(aes(label=Freq),vjust=1.6,color="white",size=4)+theme_minimal()
```
##Task 1.2 Statistics 
 step 1: Find data statistics and get minimum and maximum
```{r Summary of frequency vs Month}
uberDataMonth<-factor(uberData$month)
table(uberDataMonth)
min(table(uberDataMonth))
max(table(uberDataMonth))
```
##Task 1.2 Summary
The visualization and statistics above demonstrate the following result. April sees the lowest with a count of 564516
September sees the highest number of Uber Pickups with 1028136.

##Task 1.3
## Task 1.3 Plot
### Task 1.3 Steps
Hour is extracted from Date/Time colum . The code invlves explicit consideration of column value as Date/Time. The original csv is update with a new column knows as Hour. Generate Frequency of Hour and then Sort Data accordng to Hour. Afterwards, we p
lot bargraph with Hour on x-axis and Frequency on Y-Axis for Visualization

```{r Hour extraction}
uberData$hour <- format(strptime(uberData$'Date/Time',format = '%m/%d/%Y %H:%M:%S'), "%H")
```


```{r Plot on frequency vs Hour}
Hour_count<-as.data.frame(table(uberData$hour))
names(Hour_count)<-c("Hour","Freq")
ggplot(data=Hour_count,aes(x=Hour,y=Freq,fill=rownames(Hour_count)))+geom_bar(stat="identity",colour = "black")+guides(fill=FALSE)+labs(title = "Frequency by Hour ",x="Hour of day ",y="Frequency")+theme_minimal()
```
##Task 1.3 Statistics 
 step 1: Find data statistics and get minimum and maximum
 
```{r Summary of frequency vs hour}
uberDataHour<-factor(uberData$hour)
table(uberDataHour)
min(table(uberDataHour))
max(table(uberDataHour))
```
##Task 1.3 Summary
The visualization and statistics above demonstrate the following result
**17 hours** sees the highest number of Uber Pickups with 336190 and **02 Hours** sees the lowest with a count of 45865

## Task 2.0
### Steps
First, Obtain the New York map using get_map API and plot density map of data accroding to Latitude vs Longitude of uber pick ups. Afterwards, For better visulaization the points need to be set to size 2. Bin the points and drop the bins which dont have any samples in them

```{r Task2.0 Density Map on minor data}
register_google(key = "AIzaSyCD2-pP0yfM8aC8sjCB4g--CaCcExP6NhU")
densityPlot <- get_map("New York", zoom = 12,color="bw")
ggmap(densityPlot)+stat_density2d(aes(x = uberMinorData$Lon, y = uberMinorData$Lat,fill = ..level..),alpha=0.5,size=2,bins=30,geom="polygon",data=uberMinorData)+geom_density_2d(data=uberMinorData,aes(x = uberMinorData$Lon,y = uberMinorData$Lat),alpha=0.5,bins=30)+
  labs(title = "Uber Pickup Density By Location\n", x = "\nLongitude", y = "Latitude\n")
 
```


### Steps
Obtain the New York map using get_map API and then, plot point data wrt Latitude and Longitude


```{r Task2.0 Point Map on minor data}

pointMap <- get_map("New York", zoom = 10)
ggmap(pointMap)+geom_point(data=uberMinorData,aes(x = uberMinorData$Lon, y = uberMinorData$Lat,colour='Red'))+ scale_color_manual(values = c("Red"="red")) +  labs(title = "Uber Pickups By Location\n", x = "\nLongitude", y = "Latitude\n" ,colour="Uber Pickup")
 
```

###Summary

The above graphs(density and point plot) shows where Uber Pickups occur in NYC according to the geospatial data.

#Task 2.1

##Steps
Extract Day from DateTime column and store it as another column 
```{r Day extraction for minor data}
uberMinorData$day <- weekdays(as.Date(uberMinorData$'Date/Time',"%m/%d/%Y"))
```
#Task 2.1 Points plot
Download map for New York using get_map and set it to ny using getmap adn add points to the map using Lat and Lon data categorizing the color according to Day
```{r Task2.1 point plot minor data on Day vs frequency}
nymap <- get_map("New York", zoom = 10)
ny <- ggmap(nymap)
ny + geom_point(aes(x = Lon, y = Lat, colour = uberMinorData$day), data = uberMinorData)+labs(title = "Uber Pickups By Day on Location\n", x = "\nLongitude", y = "Latitude\n",colour="Day")
```


#Task 2.2
Extract Month from DateTime column and store it as another column 
```{r Task 2.2 Month extraction for minor data}
uberMinorData$month<- months(as.Date(uberMinorData$'Date/Time',"%m/%d/%Y"))
```
#Task 2.2 Points plot
Download map for New York using get_map and set it to ny using getmap and add points to the map using Lat and Lon data categorizing the color according to Month
```{r Task 2.2 point plot minor data on Month vs frequency}
nymap <- get_map("New York", zoom = 10)
ny <- ggmap(nymap)
ny + geom_point(aes(x = Lon, y = Lat, colour = uberMinorData$month), data = uberMinorData)+labs(title = "Uber Pickups By Month on Location\n", x = "\nLongitude", y = "Latitude\n",colour="Red")
```

#Task 2.3

Extract Hour from DateTime column and store it as another column 
```{r Task 2.3 Hour extraction of minor data}
uberMinorData$hour <- format(strptime(uberMinorData$'Date/Time',format = '%m/%d/%Y %H:%M:%S'), "%H")
```
##Steps
Download map for New York using get_map and set it to ny using getmap and add points to the map using Lat and Lon data categorizing the color according to Hour
#Task 2.3 Points plot
```{r Task 2.3 point plot minor data on Hour vs frequency}
nymap <- get_map("New York", zoom = 10)
ny <- ggmap(nymap)
ny + geom_point(aes(x = Lon, y = Lat, colour = uberMinorData$hour), data = uberMinorData)+labs(title = "Uber Pickups By Hour on Location\n", x = "\nLongitude", y = "Latitude\n",colour="Hour")
```
